### 2021-02-20 Drafts/Draft_1.tex
Draft 1 updates

 \documentclass[12pt]{article}
\usepackage[capposition=top]{floatrow}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{float}
\usepackage{eurosym}
\usepackage{csvsimple}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{caption}
\usepackage{tablefootnote}
\usepackage{ntheorem}
\usepackage{sectsty}
\usepackage[justification=centering,textfont={sc},labelfont={rm}]{caption}
\usepackage{geometry}
\geometry{margin=.5in}
\usepackage[numbers, square]{natbib}
\usepackage{graphicx}
\usepackage{setspace}
\onehalfspacing
\begin{document}
	\title{Bringing Expectations to the Collective Bargaining Table: Evidence from Brazilian Firms}
	\author{Valerie R. Boctor \\ Roberto Hsu Rocha}
	\date{\today}
	\maketitle
	\abstract{\textit{Ideal: We use matched employer-employee data and firm-level inflation expectations to examine the role of inflationary news shocks on collective bargaining outcomes in Brazil. Our quasi-natural experiment comes from firm-level variation in the timing of collective bargaining agreements (CBAs) around inflationary news shock. That is, firms with CBAs just before an inflationary news shock may have wage and employment levels that differ systematically from firms whose CBAs fall just after the shock. For firms that settled contracts just before the shock, real wages were set too high, leading to reductions in employment and investment at the firm level. Furthermore, we adapt the Diamond-Mortensen-Pissarides model to include inflation expectations to demonstrate that the macroeconomic effects of collective bargaining channel of inflationary news shocks.}}

	\section{Introduction}
	\subsection{Main Questions}
	\subsection{Economic Significance/ Literature Review}
\end{document}

### 2021-03-04 python/python_rev_2020/all/01_regressions_all.py
File organization and set-up for summary stats script

@@ -12,11 +12,11 @@ ibre_path = '/Users/Valerie/Dropbox/Courses/CoursesSpring2020/ECON236B/236B_Fina
 all_path = '/Users/Valerie/Dropbox/Courses/CoursesSpring2020/ECON236B/236B_FinalPaper/CBA monetary policy and firms/data/clean/all/'
 figures_path = '/Users/Valerie/Dropbox/Courses/CoursesSpring2020/ECON236B/236B_FinalPaper/CBA monetary policy and firms/documents/Proof_of_Concept_VB/figures/'
 
-os.chdir(all_path)
-orbis = pd.read_csv(orbis_path+'orbis10_19.csv')
-orbis = orbis.iloc[:,1:orbis.shape[1]]
-rais_firms = pd.read_csv(agreements_path+\
-	'2013_2017_firmlevel_paneldecember.csv')
+	os.chdir(all_path)
+	orbis = pd.read_csv(orbis_path+'orbis10_19.csv')
+	orbis = orbis.iloc[:,1:orbis.shape[1]]
+	rais_firms = pd.read_csv(agreements_path+\
+		'2013_2017_firmlevel_paneldecember.csv')
 rais_firms = rais_firms.iloc[:,1:rais_firms.shape[1]]
 agreements = pd.read_stata(agreements_path+\
 	'acordos_infosecnpjmerged.dta')

### 2021-04-22 Drafts/Draft_1.tex
File organization and set-up for summary stats script
Prospectus

@@
 \usepackage{float}
+\usepackage{floatrow}
 \usepackage{eurosym}
@@
 \usepackage{hyperref}
+\usepackage{adjustbox}
 \usepackage{ulem}
@@
 \usepackage{geometry}
-\geometry{margin=.5in}
+\geometry{margin=1in}
 \usepackage[numbers, square]{natbib}
@@
 \usepackage{setspace}
+\usepackage[section]{placeins}
 \onehalfspacing
@@
 	\title{Bringing Expectations to the Collective Bargaining Table: Evidence from Brazilian Firms}
-	\author{Valerie R. Boctor \\ Roberto Hsu Rocha}
+	\author{Valerie R. Boctor\thanks{This project is a collaboration with Roberto Hsu Rocha. This document, including all tables and figures, was written by Valerie Boctor. All mistakes are my own. }}
 	\date{\today}
 	\maketitle
-	\abstract{\textit{Ideal: We use matched employer-employee data and firm-level inflation expectations to examine the role of inflationary news shocks on collective bargaining outcomes in Brazil.
-Our quasi-natural experiment comes from firm-level variation in the timing of collective bargaining agreements (CBAs) around inflationary news shock.
-That is, firms with CBAs just before an inflationary news shock may have wage and employment levels that differ systematically from firms whose CBAs fall just after the shock.
-For firms that settled contracts just before the shock, real wages were set too high, leading to reductions in employment and investment at the firm level.
-Furthermore, we adapt the Diamond-Mortensen-Pissarides model to include inflation expectations to demonstrate that the macroeconomic effects of collective bargaining channel of inflationary news shocks.}}
 
+% =================================================================================================
+% 												INTRODUCTION	
+% =================================================================================================
 	\section{Introduction}
-	\subsection{Main Questions}
-	\subsection{Economic Significance/ Literature Review}
+		Using plausibly exogenous variation in the timing of collective bargaining agreements (CBAs) in Brazil, we analyze the effect of an inflationary news shock on firm-level wages and employment.
+The use of detailed micro data allows us to precisely measure labor market rigidities and the relative bargaining power of firms and workers.
+Furthermore, we assess the degree of wage rigidity for firms that negotiate just before and just after the inflationary news shock.
+This research may have important implications for both traditional and expectations-based monetary policy: In particular, we can inferences about the optimal timing of monetary policy using our precise measure of the time-varying degree of nominal wage rigidity.
+Additionally, this research may highlight the importance of the collective bargaining channel of expectations-based monetary policy.
+
+		\subsection{The Inflationary News Shock}
+			 In 2015 Q2, a combination of falling global commodity prices, fiscal malfeasance, and corruption at the state-owned oil company Petrobras sank the Brazilian economy into its deepest recession in over 30 years \cite{yukBrazilSlumpsWorst2016}, \cite{dyerBrazilEconomySlips2015}.
+Already under fire for ``creative accounting'' on costly social programs during her first term, the newly reelected President Dilma Rousseff became further embroiled when news of her involvement with the Petrobras bribery scheme emerged, solidifying public accusations of vote buying practices, and eroding Brazil's financial capital.
+Despite having been reprimanded for deferring payment on social programs during her first term,  Dilma Rousseff continued to violate Brazil's budgetary rules, leaving the Bank of Brazil to finance roughly \$4bn without formally issuing a loan \cite{phillipsChargesBrazilianPresident2016}.
+This practice, known as ``fiscal pedaling'' led to a large increase in government debt and contributed directly to growing inflation rates throughout 2015.
+Against this backdrop, consumer and business confidence plummeted, reaching their lowest level in 2015 Q3 Figure \ref{fig:BusinessConfidence}, Figure \ref{fig:ConsumerConfidence}.
+GDP contracted by more than 7\% between 2015 and 2016, while inflation rose steadily from 6.45\% in January 2015 to a high of 9.8\% in June 2016 \cite{leahyCoolHeadCrisis2017}.
+Taken together, these trends constituted a period of stagflation, not unlike the U.S.
+experience in the 1970s.
+
+			In a reversal that hearkens back to the Volcker Disinflation, Dilma Rousseff's impeachment and subsequent replacement by Acting President Michael Temer led to significant changes in monetary policy that effectively curbed inflation and set the Brazilian economy on the path to recovery.
+Michel Temer appointed technocrat Ilan Goldfajn as the Brazilian Central Bank.
+Goldfajn's credible inflation targeting, interest rate cuts, and reassurance on improved fiscal responsibility helped to stabilize the economy and reduce inflation to under 3.5\% by the close of 2017.
+
+			 We use anticipation of the regime shift as an inflationary news shock.
+As shown in Figure \ref{fig:Inflation} below, inflation expectations peaked in February 2016 and decline sharply.
+We attribute the inflection point to a wave of public protests that erupted in March 2016: during that month, a leaked phone-call between Rousseff and her disgraced predecessor, former President Luiz Inácio de Silva, revealed her attempts to shield him from prosecution related to the Petrobras scandal.
+These protests represented a final push to impeach and a public signal that Rousseff's government would be overturned \cite{douglasReleaseTappedPhone2016}, Figure \ref{fig:Inflation}.
+It is important to note that speculation of Rousseff's impeachment and potential replacement by Michel Temer (and Ilan Goldfajn) began to appear in the news as early as December 2015, and this messaging was credible and consistent over time.
+Therefore, it is reasonable to assume the public equated Rousseff's removal from office with the advent of Temer and Goldfajn and an associated reduction in future inflation.
+
+			\begin{figure}[!ht]
+			\centering
+			\includegraphics[scale = .5]{tables-figures/Realized_Inflation_and_Consumer_Expectations.png}
+			\caption{The inflationary news shock}
+			\label{fig:Inflation}
+			\floatfoot{Consumer inflation expectations peak in February 2016 and begin to decline thereafter.
+This timing corresponds to the March 2016, followed by Rousseff's impeachment in April, and suspension of powers in July of the same year.}
+			\end{figure}
+
+			\begin{figure}[!ht]
+				\includegraphics[scale = .8]{old_figures/IBRE_Consumer_Expectations.png}
+				\caption{Consumer Confidence Index}
+				\floatfoot{Consumer confidence dropped precipitously during the political-economic crisis and began to recover as speculation of Rousseff's replacement by Temer's administration took root.}
+				\label{fig:ConsumerConfidence}
+			\end{figure}
+
+			\begin{figure}[!ht]
+				\includegraphics[scale = .85]{old_figures/IBRE_Firm_Expectations.png}
+				\caption{Business Confidence Index}
+				\label{fig:BusinessConfidence}
+				\floatfoot{Business confidence trends across all major sectors strong comovement with consumer confidence, shown in Figure \ref{fig:ConsumerConfidence}.}
+			\end{figure}
+
+			\FloatBarrier
+		\subsection{Main Research Questions}
+		In this project, we are mainly interested in estimating the effects of inflationary news shocks on collective bargaining outcomes in Brazil and understanding macroeconomic implications of these effects.
+We summarize the main quesstions we will address below:
+		\begin{enumerate}
+			\item How does the timing of CBAs affect the response of wages and employment to the inflationary news shock?
+
+			\item To what extent do firm-level wages exhibit nominal rigidities?
+
+			\item What are the implications of wage rigidities and employment behavior for optimal monetary policy?
+		\end{enumerate}
+% =================================================================================================
+% 												BACKGROUND	
+% =================================================================================================
+	\section{Background}
+	In this section, we discuss our intended contributions to the literature and provide key information about collective bargaining institutions in Brazil.
+		\subsection{Literature Review}
+		Olivei and Tenreyro (2007) posit that monetary policy shocks in the US have larger and less persistent effects during the first and second quarters than in the latter half of the year do to the timing of wage contract negotiations \cite{oliveiTimingMonetaryPolicy2007}.
+Because negotiations are more likely to occur in the first two quarters in the US, wages are more flexible during this period and respond more readily to monetary policy shocks.
+Since wages are less likely to be renegotiated during the third and fourth quarters, the response to monetary policy shocks during these periods tends to be smaller and more protracted.
+Our research builds on Olivei and Tenreyro's findings by measuring precisely the effect of an aggregate shock on firm-level wages, taking into account the relative timing of each firm's most recent CBA.
+If the reduction in nominal wages and/or employment is relatviely larger and less persistent for firms CBAs just after the news shock, then our results would corroborate Olivei and Tenreyo's aggregate findings.\footnote{See ``CBAs in Brazil'' for discussion of legally mandated downward nominal wage rigidity and workarounds.}
+
+		\begin{itemize}
+			\item Discuss David Card findings on unions and wage outcomes \cite{cardUnexpectedInflationReal1990}, \cite{cardEffectUnionsStructure1996}, \cite{cardStrikesBargainingSurvey1990}
+			\item Discuss relevance to literature on nominal rigidities and macro models \cite{taylorStayingPowerStaggered2016} \cite{taylorChapter15Staggered1999} \cite{fukuiTheoryWageRigidity}
+		\end{itemize}
+
+		\subsection{Collective Bargaining Agreements in Brazil}
+		In Brazil, collective bargaining agreements typically last 1 or 2 years.
+All CBA contracts have a maximum duration of 2 years, but can be extended for additional years.
+\cite{lagosLaborMarketInstitutions2021}.
+CBAs typically set a fixed nominal wage levels for the duration of the contract.
+It is important to note that nominal wages are legally required to be downward rigid.
+That is, nominal wages cannot be reduced unless the union approves the reduction first.
+Without an affirmation from the workers' union, if nominal wages fall, firms are legally required to offer workers additional hours.
+Conversely, a reduction in workers' hours must be compensated by a commensurate increase in the nominal wage rate.
+\cite{EnglishTranslationConsolidation2010}
+		Given the constraints facing firms, wages (measured in average monthly remuneration) may still be reduced through separations, or by reclassifying workers to occupations with lower nominal wage levels.
+
+		In our context, a natural question that arises is whether the frequency and duration of CBAs may be endogenous to inflationary news shocks or changes in economic uncertainty.
+Below, we present evidence that the timing of CBAs is largely consistent over the entire sample.
+	\begin{figure}[!ht]
+		\centering
+		\caption{CBA Hazard}
+		\label{fig:CBAHazard}	
+		\includegraphics[scale = .72]{tables-figures/cba_hazard.png}
+		\floatfoot{The hazard rate for both the pre- and post- shock groups peaks at 12 months.}
+		\end{figure}
+		Figure \ref{fig:CBAHazard} shows the CBA hazard rate for firms with CBAs before and after the March 2016 shock.
+For both groups, the hazard rate peaks at 12 months, which suggests that the duration of contracts does not shrink in response to possible uncertainty about future inflation.
+
+		\begin{figure}[!ht]
+			\centering
+			\caption{Probability of CBA each period}
+			\label{fig:CBAProb}
+			\includegraphics[scale = .5]{tables-figures/cba_probability.png}
+			\floatfoot{For each year in the sample, the two most common months for CBAs to occur are January and May. We do not observe changes in the distribution during the 2015-2016 crisis.}
+		\end{figure}
+
+		Figure \ref{fig:CBAProb} shows the probability of having a CBA at each period in the sample.
+Note that the probability of having a CBA does not increase in 2015-2016.
+This graph suggests that CBAs do not become more frequent in response to news shocks or uncertainty about future inflation.
+		Combining information from Figure \ref{fig:CBAHazard} and Figure \ref{fig:CBAProb}, we conclude that the timing of CBAs is plausibly exogenous.
+
+		Furthermore, we can rule out the possibility that the distribution of CBAs over time is driven by any particular industry.
+As Lagos (2021) points out, there is no bunching by certain industries around any particular month for CBAs \cite{lagosLaborMarketInstitutions2021}
+% =================================================================================================
+% 													DATA	
+% =================================================================================================
+	\section{Data}
+		\subsection{RAIS}
+		The Annual Social Information Report (RAIS) is an administrative dataset containing detailed information about workers and employers at the monthly frequency.
+The key variables of interest are monthly remuneration, employment status, 6-digit occupation codes (CBO), and 2-digit broad sector classifications from IBGE.
+We use average monthly remuneration by firm as the main wage variable throughout the study.
+We use the count of employed workers in a given month as the main employment variable of interest.
+		\begin{figure}
+			\centering
+			\caption{INSERT RAIS SUM STATS}
+			\label{tab:SumStatsRAIS}
+		\end{figure}
+
+		\begin{figure}
+			\centering
+			\caption{INSERT RAIS RAW SAMPLE}
+			label{tab:RawDataRAIS}
+		\end{figure}
+
+		\subsection{Sistema Mediador}
+		Sistema Mediador is the Brazilian Ministry of Labor's online database of all CBAs.
+We use the subset of collective bargaining agreements that contain individual firm-worker agreements (acordos).
+This data is webscraped from the Sistema Mediador website.
+The key variables we use for this study the CBA start and end dates, and company names, which we use to link the CBA data with RAIS.
+		\begin{figure}
+			\centering
+			\caption{INSERT SM SUM STATS}
+			\label{tab:SumStatsRAIS}
+		\end{figure}
+
+		\begin{figure}
+			\centering
+			\caption{INSERT SM RAW SAMPLE}
+			label{tab:RawDataRAIS}
+		\end{figure} 
+		\subsection{FGV IBGE}
+		We use aggregate consumer inflation expectations fron the Getuglio Vargas Foundation (FGV).
+In the future, we may gain access to firm-specific expectations of their own prices and costs.
+This information would alllow us to estimate the relationship between expectations and wage adjustments more precisely.
+		\begin{figure}
+			\centering
+			\caption{INSERT FGV SUM STATS}
+			\label{tab:SumStatsRAIS}
+		\end{figure}
+
+		\begin{figure}
+			\centering
+			\caption{INSERT FGV RAW SAMPLE}
+			label{tab:RawDataRAIS}
+		\end{figure}
+
+% =================================================================================================
+% 											EMPIRICAL STRATEGY	
+% =================================================================================================
+	\section{Research Design}
+		\subsection{Reduced Form Model}
+		We use the March 16 protests and anticipation of Rousseff's impeachment as the inflationary news shock.
+We estimate the effects on wages over time for firms in the pre- and post- shock groups.
+We restrict our sample to the subset of firms with contracts of at least 12 months durations beginning either in January 2016 or May 2016.
+We use January and May 2016 as our pre- and post-shock dates, respectively for two reasons: 1.) Average inflation expectations differ substantially across these periods\footnote{See Figure \ref{fig:Inflation}}, and 2.) most CBAs in a given year occur during these months.\footnote{See Figure \ref{fig:CBAProb}} This allows us to take advantage of a substantial part of the variation in the data.
+	\begin{align}
+		\intertext{We write down the reduced form model for wages as:} \label{eqn:rem_DiD}
+		\log{w_{i,t}} & = \alpha_i + \delta_t + \sum_{t=\text{Jan 2015}, t\neq t^{\text{shock}}}^{\text{Dec 2017}} \beta_{t}D_{i,t} + \varepsilon_{i,t} 
+		\intertext{Similarly, the model for number of workers is:}  \label{eqn:emp_Did}
+		\log{n_{i,t}} & = \alpha_i + \delta_t + \sum_{t=\text{Jan 2015}, t\neq t^{\text{shock}}}^{\text{Dec 2017}} \beta_{t}D_{i,t} + \varepsilon_{i,t} 
+	\end{align}
+
+	In this specification, $D_{i,t}$ is a dummy variable that equals 1 if firm $i$ is in the post-shock group and if the time period of the observation is equal to $t$.
+A negative estimate for $\beta_t$ would suggest that at time period t, the log wages for firms in the post-shock group fell by more than the pre-shock group.
+$\alpha_i$ is a firm-level fixed effect, and $\delta_t$ is a time fixed effect.
+
+	If wages were perfectly flexible, our prediction, in line with Olivei and Tenreyro (2007), would be that the $\beta_t$s would be most negative just after the shock and attenuate to zero over time.
+However, given the presence of legally required downward nominal rigidity, firms are left with only indirect ways to reduce workers' monthly earnings.
+For example, monthly earnings would fall if firms separate from workers or reclassify them to occupations with lower implied wages.
+Note that simply reducing hours would not result in a reduction in monthly earnings, since the Brazilian Consolidation of Labor Laws (CLT) requires that hourly reductions be compensated with a commensurate increase in nominal wage rates.
+Moreover, because we restrict our analysis to a balanced panel, nominal wage reductions cannot be explained by firm exit/entry.
+
+		\subsection{Preliminary Results}
+		\begin{figure}[!ht]
+			\centering
+			\caption{Monthly Earnings per Worker}
+			\label{fig:NormWages}
+			\includegraphics[scale = .65]{tables-figures/normalized_avg_med_rem_2015_2017.png}
+			\floatfoot{Each series is normalized by the values by December 2015 values.
+The sample is restricted to a balanced panel of firms between 2015 - 2017 with contracts beginning either in January or May 2016 and lasting at least 12 months.}
+		\end{figure}
+		One of the most striking features of this diagram is the sharp drop in all measures monthly nominal earnings in between December 2015 and January 2016.
+Surely, part of this drop is due to seasonality: Since bonuses are typically paid out in December, monthly earnings temporarily rise in December and fall back to the trend thereafter.
+(See November 2016 - January 2017 for comparison).
+However, the large drop in monthly earnings between December 2015 and January 2016 is salient because wages fall to levels far below the pre-trend levels.
+While the timing of the drop does not line up with our ``shock'' in March 2016, this graph demonstrates the magnitude of the recession vis-à-vis workers' earnings.
+We discuss the difference in nominal wage growth between the pre- and post-shock groups in the regression results in Figure \ref{fig:RegressionResults}.
+
+		\begin{figure}[!ht]
+			\centering
+			\caption{Number of Workers}
+			\label{fig:NormEmp}
+			\includegraphics[scale = .65]{tables-figures/normalized_avg_med_emp_2015_2017.png}
+			\floatfoot{Each series is normalized by the values by December 2015 values.
+The sample is restricted to a balanced panel of firms between 2015 - 2017 with contracts beginning either in January or May 2016 and lasting at least 12 months.}
+		\end{figure}
+
+		In Figure \ref{fig:NormEmp}, we note that the average and median number of workers per firm is falling during the sample period.
+This graph provides preliminary evidence that part of the reduction in monthly earnings way be due to separations.
+
+		Figure \ref{fig:RegressionResults} plots the $\beta_t$ estimates from equations \ref{eqn:emp_Did} and \ref{eqn:rem_DiD} above.
+Each point represents the difference in nominal wage growth between firms that negotiated just before the March 2016 shock ($t=0$) and just after.
+We restrict the sample to a balanced panel of firms between 2015 and 2016 whose contracts begin either in January 2016 or May 2016 and last at least 12 months.
+		\begin{figure}[!ht]
+			\centering
+			\caption{Response of Wages and Employment Given CBA Timing}
+			\includegraphics[scale = .49]{tables-figures/DiD_Plots_Weighted_Employment_Rem.png}
+			\label{fig:RegressionResults}
+			\floatfoot{Regression results from DiD estimation in equations \ref{eqn:emp_Did} and \ref{eqn:rem_DiD}.
+The results for the wage regressions use analytic weights based on the number of workers at firm $i$ in period $t$.
+}
+		\end{figure}
+		Contrary to our initial prediction, the DiD results suggest that firms that negotiated contracts just after disinflationary shock had higher nominal wage levels than their counterparts.
+Furthermore, this result is not accompanied by a relative reduction in the number of workers for the post-shock group.
+Instead, firms in the post-shock exhibit both relatively high nominal wages and employment levels, despite the fact that the real cost to group suggest that the difference in nominal wage growth is statistically indistinguishable from zero until January 2015 (t= -14), represented by the red dashed line on the graph.
+The coefficients begin to fluctuate after January 2015, as economic uncertainty mounted.
+The fluctuations become even more pronounced after March 2016 (t=0).
+
+	Tables \ref{tab:wage_reg} and \ref{tab:emp_reg} show several variants of the specifications displayed in Figure \ref{fig:RegressionResults}. 
+		\begin{figure}[!ht]
+				\caption{Log Wage Regressions}
+				\label{tab:wage_reg}
+				\tiny{\input{tables-figures/DiD_Rem_Results}}
+				\floatfoot{B/ NB = Balanced/ Unbalanced, W/ NW = Weighted/ Unweighted. \\ Weights are based on number of workers at firm $i$ in period $t$.}
+		\end{figure}
+
+		\begin{figure}[!ht]
+			\caption{Log Employment Regressions}
+			\label{tab:emp_reg}
+			\tiny{\input{tables-figures/DiD_Emp_Results}}
+			\floatfoot{B/ NB = Balanced/ Unbalanced}
+		\end{figure}
+
+% =================================================================================================
+% 													THEORY	
+% =================================================================================================
+	\section{Future Research}
+	Further analysis is required to pin down what factors are driving the reduction in nominal wages across all firms during the 2015-2016 recession.
+\ref{fig:NormWages} One possible explanation is that firms separate from workers during the recession, curtailing the workers' monthly earnings.
+There is some preliminary evidence of this in Figure \ref{fig:NormEmp}.
+Another potentially complementary explanation is that firms reclassify workers to occupations with lower implied nominal wages.
+While we have not yet carried out the analyses to corroborate whether the latter explanation is pertinent, we may have sufficient data to accomplish this in the near future.
+
+	Another salient question is why firms that negotiate wages just after the shock set higher nominal wages than firms negotiating just before.
+\ref{fig:RegressionResults} One possible explanation is that firms in the post-shock group have higher levels of optimism about the future state of the economy and their own future profit margins, which may lead them to set higher wage levels.
+We may gain access to microdata on firms' own-price and own-cost expectations that would allow us to study the link between individual firms' expectations and negotiated wages as well as hiring and firing decisions.
+
+	Following Olivei and Tenreyo (2007), we may adapt a dynamic stochastic general equilibrium model with uneven staggering of wage contracts and downward rigidity. \cite{oliveiTimingMonetaryPolicy2007}
+
+	Another approach would be to adapt a DMP model to include inflation expectations, similar in spirit to Fukui (2020).  \cite{fukuiTheoryWageRigidity}
+
+% =================================================================================================
+% 												FUTURE RESEARCH	
+% =================================================================================================
+
+	\newpage
+	\bibliographystyle{chicago}
+	\bibliography{Bibliography-CBAs-Inflation}
 	\end{document}

### 2021-04-22 Drafts/QE Docs/Summary_03.05.21.tex
Prospectus

 \documentclass[12pt]{article}
\usepackage[capposition=top]{floatrow}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{float}
\usepackage{eurosym}
\usepackage{csvsimple}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{caption}
\usepackage{tablefootnote}
\usepackage{ntheorem}
\usepackage{sectsty}
\usepackage[justification=centering,textfont={sc},labelfont={rm}]{caption}
\usepackage{geometry}
\geometry{margin=.25in}
\usepackage[numbers, square]{natbib}
\usepackage{graphicx}
\usepackage{setspace}
\onehalfspacing
\begin{document}
	\title{\small{Bringing Expectations to the Collective Bargaining Table: Evidence from Brazilian Firms}}
	\author{\small{Valerie R. Boctor}\thanks{This project is a collaboration with Roberto Hsu Rocha. This document was written by Valerie Boctor. All mistakes are my own.}}
	\date{\vspace{-1.5cm}}
	\maketitle
	\abstract{We use matched-employer and employee data from Relacao Anual de Informacoes Socials (RAIS) as well as information on collective bargaining outcomes from Sistemas Mediador to study the effects of a major inflationary news shock on collective bargaining outcomes and macroeconomic implications in Brazil. A major news shock occurred during the impeachment of former President Dilma Roussef in early 2016: Inflation expectations plummeted and stabilized from 11.5\% in February 2016 to around 3\% by May 2017. Using variation in the timing of collective bargaining agreements at the firm level, we assess the impact of inflation expectations on firms' wages and employment. Furthermore, we examine the extent to which downward nominal wage rigidity binds at the firm level, as well as alternative avenues firms may use to compensate for suboptimal real wages. We will use a we will use an adapted staggered pricing model assess the macroeconomic implications of collective bargaining activity driven by inflationary news shocks. 
	\begin{center}\textbf{Data} \end{center}
	In addition to publicly available aggregate inflation expectations from the Brazilian Central Bank, we use two main data sets to carry out this study: RAIS, which contains highly detailed information about individual workers, employers and unions, as well as the collective bargaining data from Sistemas Mediador. Using the latter database, we are able to obtain detailed information about firm-to-worker collective bargaining agreements, including bargained wages, effective dates, and firm IDs. While this data is arguably sufficient to answer the main research questions, we are also in the process of obtaining firm-level expectations of their own prices and costs. This data would be useful in the way of establishing an empirical relationship between firms' price expectations and resultant wages. This information may also help us to assess whether firms are behaving in a manner consistent with a staggered pricing scheme.

	\begin{center} \textbf{Economic Motivation} \end{center}
	This project is motivated by the sheer size of the drop in inflation expectations around the Roussef's impeachment and anticipation of a major regime shift. This event can be viewed as a ``modern-day Volcker shock,'' providing ample variation to study the effects on collective bargaining activity using unique, highly detailed micro data. This project also has implications for our understanding of forward guidance, which is typically thought to operate through consumption and investment. Our project adds to these by establishing empirical evidence that the collective bargaining channel may also play an important role in macroeconomic performance. Our research also provides an empirical test for Tenreyro \& Olivei's (2007 AER) conjecture that the timing of monetary policy shocks in the US matters due to the seasonality in wage rigidity, which is ostensibly driven by the timing of wage contracts. 
		\begin{center} \textbf{Committee Members} \end{center} Yuriy Gorodnichenko (Dissertation Advisor), Emi Nakamura (tentative Committee Chair), Benjamin Schoefer, David Card
\end{document}

### 2021-04-22 Drafts/VBoctor_Slides April_2021.tex
Prospectus

\documentclass{beamer}
\usetheme{default}
\setbeamertemplate{footline}[frame number]
\title{Bringing Expectations to the Collective Bargaining Table}
\subtitle{Evidence From Brazilian Firms}
\date{\today}
\usepackage{booktabs}
\begin{document}
\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\section{Introduction}
\begin{frame}
\frametitle{Main Questions}
	\begin{itemize}
		\item How do inflationary news shocks affect collective bargaining outcomes (wages and employment)?
		\item How does the timing of collective bargaining agreements (CBAs) affect firm performance?
		\item What are the macroeconomic implications of collective bargaining activity induced by inflationary news shocks? 
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Economic Significance}
	\begin{itemize}
		\item Inflation expectations and economic behavior
		\item Monetary policy application: forward guidance
			\begin{itemize}
				\item Traditional channels are consumption and investment. Analyze potential CBA channel
			\end{itemize}
		\item Timing of monetary policy Shocks (Tenreyro \& Olivei 2007 AER)
		\item Nominal rigidity \--- wage adjustment asymmetries (Kaur 2019 AER)
	\end{itemize}
\end{frame}

\section{Data}
\begin{frame}
\frametitle{Data}
	\begin{itemize}
		\item RAIS \--- Matched employer-employee data 
		\item Sistemas Mediador \--- collective bargaining agreements (single firm - employee contracts)
		\item IBRE / IBGE \--- aggregate inflation expectations	(May gain access to firm-level price and cost expectations)
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Background}
	\begin{itemize}
		\item Dual economic and political crises in Brazil 2015 - 2016
		\item  Impeachment of President Dilma Rousseff announced on December 2, 2015. Her powers were suspended on May 12, 2016.
		\item \textbf{Inflationary news shock:} In June 2016, acting President Michel Temer appointed Ilan Goldfajn as the BCB head.
		\centering
		\includegraphics[scale=.2]{old_figures/goldfajn.png}
	\end{itemize}
\end{frame}

\section{Preliminary Findings}
\begin{frame}
\frametitle{Preliminary Findings}
	\begin{itemize}
	\item Across all sectors, inflation expectations rose to 12\% until February 2016, at which point they began to decline to roughly 5\% by Q3 2017 (as anticipation of new regime took root.)
	\item Timing of CBAs provides opportune variation for studying effects on wages, employment
	\item Setting may allow us to analyze multiple inflation shocks prior to 2017
	\end{itemize}
\end{frame} 
\begin{frame}
		\centering
		\includegraphics[scale=.6]{old_figures/IBRE_Consumer_Inflation_Expectations.png}
\end{frame}

\begin{frame}
		\centering
		\includegraphics[scale=.6]{old_figures/IBRE_Consumer_Expectations}
\end{frame}

\begin{frame}
		\centering
		\includegraphics[scale=.6]{old_figures/IBRE_Firm_Expectations}
\end{frame}

\begin{frame}
\frametitle{Actual Inflation in Brazil}
	\centering
	\includegraphics[scale=.45]{old_figures/Brazil_Inflation.png}
\end{frame}

\begin{frame}
\frametitle{Average Monthly Real Wages in Brazil}
	\centering
	\includegraphics[scale=.35]{old_figures/Wage_Growth.png}
\end{frame}

\begin{frame}
\frametitle{Timing of CBAs in Brazil}
		\begin{center}
		\includegraphics[scale=.45]{tables-figures/cba_hazard.png}
		\end{center}
\end{frame}

\begin{frame}
\frametitle{Timing of CBAs in Brazil}
		\centering
		\includegraphics[scale=.3]{tables-figures/cba_probability.png}
\end{frame}

\begin{frame}
	\frametitle{Reduced-Form Model 1}
	We focus on the subset of firms with a one-year contract set either in Jan 2016 (pre-shock) or Mar 2016 (post-shock). Using a DiD approach, we measure the effect on nominal wage growth of setting wages just before the shock and just after:
	\begin{align*}
		nom\:wage_{i,t} = \beta_0 + \beta_1 D_{t}^{March\:2016} + \beta_2 D^{March\:CBA}_{i} +\\
		 \beta_3[ D_{t}^{March\:2016} \times D^{March\:CBA}_{i}] + \mathbf{Z}_{i,t} + \varepsilon_{i,t}
 	\end{align*}
 	\begin{itemize}
 		\item $\beta_3$ captures the difference in nominal wage differences across firms with pre- and post- shock CBAs
 		\item Caveat: nominal wage growth may be driven by other forms of negotiation, even if a firm is ``locked-into''  a CBA
 	\end{itemize}
\end{frame}
\begin{frame}
	\frametitle{Reduced-Form Model 2}
		A preliminary reduced-form model takes the following form:
		\begin{align*}
			nominal\_wage\_growth_{i,t} & = \alpha + \beta_1 exp\_inflation_t \times CBA\_month_{i,t} +\\
			& \beta_2 CBA\_month_{i,t} + \beta_3 exp\_inflation_t + \mathbf{Z_{i,t}} + \varepsilon_{i,t} 	
		\end{align*}
\end{frame}

\begin{frame}
	\begin{center}
			\scalebox{.45}{\input{old_figures/update2_regression2.tex}}
	\end{center}
\end{frame}

\begin{frame}
\frametitle{Structural Identification Strategy}
	\begin{itemize}
		\item Adapt a staggered wage model with collective bargaining to assess impacts of monetary policy  
		\item Estimate effects on economy-wide wages, levels of employment
	\end{itemize}
\end{frame}
\end{document}

### 2021-04-22 python/python_rev_2020/IBRE/02_figures_IBRE.py

@@ -6,13 +6,25 @@ import os
 from matplotlib import pyplot as plt
 from datetime import datetime, timedelta
 from collections import OrderedDict
+# Connection to FRED
+import os
+from fredapi import Fred
+
+os.environ['api_key'] = 'fd13f48c5556bfa5a8fb4e49e90876d3'
+key =os.getenv('api_key')
+fred = Fred(key)
+actual_inflation = fred.get_series('CPALTT01BRM659N')
+actual_inflation = actual_inflation.rolling(12).mean() #rolling average of 12 months ahead (monthly) inflation
+#actual_inflation = actual_inflation.pct_change(periods = 12)
+actual_inflation = actual_inflation.loc['2005-09-01':'2020-03-01']
 
 pd.set_option('display.max_columns', 500)
 pd.set_option('display.width', 1000)
 pd.set_option('display.float_format', lambda x: '%.2f' % x)
 
-figures_path = '/Users/Valerie/Dropbox/Courses/CoursesSpring2020/ECON236B/236B_FinalPaper/CBA monetary policy and firms/documents/Proof_of_Concept_VB/figures/'
-ibre_clean_path = '/Users/Valerie/Dropbox/Courses/CoursesSpring2020/ECON236B/236B_FinalPaper/CBA monetary policy and firms/data/clean/IBRE/'
+figures_path = '/Users/Valerie/CBAs-Inflation-Git/drafts/old_figures'
+new_figures_path = '/Users/Valerie/CBAs-Inflation-Git/drafts/tables-figures/'
+ibre_clean_path = '/Users/Valerie/Dropbox/Research/CBA monetary policy and firms/data/clean/IBRE/'
 os.chdir(ibre_clean_path)
 
 # ~~~~~~~~~~~~~~~~~~~~~~ Summary statistics ~~~~~~~~~~~~~~~~~~~~~
@@ -64,11 +76,21 @@ plt.xticks(np.arange(0, len(t), 7), rotation = 45)
 plt.title('Consumer Confidence')
 plt.savefig(figures_path+'IBRE_Consumer_Expectations.png')
 fig.clear()
-# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~ Consumer Inflation Expectations ~~~~~~~~~~~~~~~~~~~~~~
+# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~ Realized Inflation & Consumer Inflation Expectations ~~~~~~~~~~~~~~~~~~~~~~
 exp_inflation = np.array(df2.exp_inflation)
[…]

### 2021-04-22 python/python_rev_2021/01_summary_stats.py

@@ -4,6 +4,7 @@ import pandas as pd
 import re
 from matplotlib import pyplot as plt
 import os
+import datetime
 
 agreements_path = '/Users/Valerie/Dropbox/Research/CBA monetary policy and firms/data/clean'
 os.chdir(agreements_path)
@@ -15,93 +16,141 @@ year = graph_df.year.astype(int)
 monthly_date = 100*year + month
 graph_df['monthly_date'] = monthly_date
 
+date_vars = ['monthly_date']
+for i in date_vars:
+	graph_df[i] = pd.to_datetime(graph_df[i], yearfirst='True', format = '%Y%m')
 #check number of non NaN observations:
 # np.sum(cba_df.COLUMN_NAME.count())
 
 # make a dummy variable that equals the sum of the month of contract and the contract duration is >= 12
-graph_df = graph_df[['identificad', 'monthly_date', 'rem','duration', 'cba_month']]
+graph_df = graph_df[['identificad', 'monthly_date', 'rem','duration', 'cba_month', 'num_workers']]
 graph_df['cba_time'] = graph_df.monthly_date[graph_df.cba_month ==1]
-cba_jan16 = (graph_df.cba_time == 201601).astype(int)
-cba_mar16 = (graph_df.cba_time== 201603).astype(int)
+cba_jan16 = (graph_df.cba_time == '2016-01').astype('int')
+cba_may16 = (graph_df.cba_time== '2016-05').astype('int')
 
 N = graph_df.identificad.count() #this is the number of rows in the data set before merging, including rows with NaN values
 graph_df['cba_jan16'] = cba_jan16
-graph_df['cba_mar16'] = cba_mar16
+graph_df['cba_may16'] = cba_may16
 
-date_vars = ['monthly_date']
-for i in date_vars:
-	graph_df[i] = pd.to_datetime(graph_df[i], yearfirst='True', format = '%Y%m')
 
-# make dummy variables at the firm level which tell if firms had cbas in Jan 2016, Mar 2016.
-#=======================================================================================================================
+# make dummy variables at the firm level which tell if firms had cbas in Jan 2016, may 2016.
[…]

### 2021-04-22 python/python_rev_2021/02_clean_data.py

@@ -1,0 +1,17 @@
+# I have no idea what I'm doing, but here goes nothing!
+from __future__ import division
+import numpy as np
+import pandas as pd
+import re
+from matplotlib import pyplot as plt
+import os
+
+data_path = '/Users/Valerie/Dropbox/Research/CBA monetary policy and firms/data/clean/RAIS/deidentified_RAIS'
+os.chdir(data_path)
+# Read in the masked raw data
+# Need to generate the wage series for only incumbent workers
+# cry sob
+
+df15 = pd.read_csv('workers_bargainedfirms2015.csv')
+df16 = pd.read_csv('workers_bargainedfirms2016.csv')
+df17 = pd.read_csv('workers_bargainedfirms2017.csv')

### 2021-04-22 Drafts/Draft_1.tex
Citations edit

@@
 \geometry{margin=1in}
-\usepackage[numbers, square]{natbib}
+\usepackage{natbib}
 \usepackage{graphicx}
@@
 		\subsection{Literature Review}
-		Olivei and Tenreyro (2007) posit that monetary policy shocks in the US have larger and less persistent effects during the first and second quarters than in the latter half of the year do to the timing of wage contract negotiations \cite{oliveiTimingMonetaryPolicy2007}.
+		Olivei and Tenreyro (2007) posit that monetary policy shocks in the US have larger and less persistent effects during the first and second quarters than in the latter half of the year do to the timing of wage contract negotiations.
 Because negotiations are more likely to occur in the first two quarters in the US, wages are more flexible during this period and respond more readily to monetary policy shocks.
@@
 
-		\begin{itemize}
-			\item Discuss David Card findings on unions and wage outcomes \cite{cardUnexpectedInflationReal1990}, \cite{cardEffectUnionsStructure1996}, \cite{cardStrikesBargainingSurvey1990}
-			\item Discuss relevance to literature on nominal rigidities and macro models \cite{taylorStayingPowerStaggered2016} \cite{taylorChapter15Staggered1999} \cite{fukuiTheoryWageRigidity}
-		\end{itemize}
+		% \begin{itemize}
+		% 	\item Discuss David Card findings on unions and wage outcomes \cite{cardUnexpectedInflationReal1990}, \cite{cardEffectUnionsStructure1996}, \cite{cardStrikesBargainingSurvey1990}
+		% 	\item Discuss relevance to literature on nominal rigidities and macro models \cite{taylorStayingPowerStaggered2016} \cite{taylorChapter15Staggered1999} \cite{fukuiTheoryWageRigidity}
+		% \end{itemize}
 
@@
 All CBA contracts have a maximum duration of 2 years, but can be extended for additional years.
-\cite{lagosLaborMarketInstitutions2021}.
-CBAs typically set a fixed nominal wage levels for the duration of the contract.
+\cite{lagosLaborMarketInstitutions2021} CBAs typically set a fixed nominal wage levels for the duration of the contract.
 It is important to note that nominal wages are legally required to be downward rigid.
@@
 		Furthermore, we can rule out the possibility that the distribution of CBAs over time is driven by any particular industry.
-As Lagos (2021) points out, there is no bunching by certain industries around any particular month for CBAs \cite{lagosLaborMarketInstitutions2021}
+As Lagos (2021) points out, there is no bunching by certain industries around any particular month for CBAs.
 % =================================================================================================
@@
 	\section{Future Research}
-	Further analysis is required to pin down what factors are driving the reduction in nominal wages across all firms during the 2015-2016 recession.
-\ref{fig:NormWages} One possible explanation is that firms separate from workers during the recession, curtailing the workers' monthly earnings.
+	Further analysis is required to pin down what factors are driving the reduction in nominal wages across all firms during the 2015-2016 recession.\footnote{See Figure \ref{fig:NormWages}} One possible explanation is that firms separate from workers during the recession, curtailing the workers' monthly earnings.
 There is some preliminary evidence of this in Figure \ref{fig:NormEmp}.
@@
 
-	Another salient question is why firms that negotiate wages just after the shock set higher nominal wages than firms negotiating just before.
-\ref{fig:RegressionResults} One possible explanation is that firms in the post-shock group have higher levels of optimism about the future state of the economy and their own future profit margins, which may lead them to set higher wage levels.
+	Another salient question is why firms that negotiate wages just after the shock set higher nominal wages than firms negotiating just before Figure \ref{fig:RegressionResults} One possible explanation is that firms in the post-shock group have higher levels of optimism about the future state of the economy and their own future profit margins, which may lead them to set higher wage levels.
 We may gain access to microdata on firms' own-price and own-cost expectations that would allow us to study the link between individual firms' expectations and negotiated wages as well as hiring and firing decisions.
 
-	Following Olivei and Tenreyo (2007), we may adapt a dynamic stochastic general equilibrium model with uneven staggering of wage contracts and downward rigidity. \cite{oliveiTimingMonetaryPolicy2007}
+	Following Olivei and Tenreyo (2007), we may adapt a dynamic stochastic general equilibrium model with uneven staggering of wage contracts and downward rigidity.
 
-	Another approach would be to adapt a DMP model to include inflation expectations, similar in spirit to Fukui (2020).  \cite{fukuiTheoryWageRigidity}
+	Another approach would be to adapt a DMP model to include inflation expectations. \cite{fukuiTheoryWageRigidity}
 

### 2021-04-22 python/python_rev_2021/01_summary_stats.py

@@ -126,6 +126,7 @@ plt.plot(emp_cba_may16_med.monthly_date, emp_cba_may16_med.emp_cba_may16_med, la
 plt.plot(emp_cba_jan16_med.monthly_date, emp_cba_jan16_med.emp_cba_jan16_med, label = "Median -- CBAs in Jan 2016")  
 plt.plot(med_emp.monthly_date, med_emp.num_workers, label = "Median -- All Firms in Sample" )
 plt.title("Workers per Firm (Monthly) 2015-2017")
+plt.axvline(x = '2016-02', color = 'black', linestyle='-')
 plt.legend()
 plt.savefig('/Users/Valerie/CBAs-Inflation-Git/drafts/tables-figures/avg_med_emp_2015_2017.png')
 plt.clf()
@@ -139,6 +140,7 @@ plt.plot(rem_cba_jan16_med.monthly_date, rem_cba_jan16_med.rem_cba_jan16_med, la
 plt.plot(med_rem.monthly_date, med_rem.rem, label = "Median -- All Firms in Sample" )
 plt.title("Monthly Remuneration 2015-2017")
 plt.legend()
+plt.axvline(x = '2016-02', color = 'black', linestyle='-')
 plt.savefig('/Users/Valerie/CBAs-Inflation-Git/drafts/tables-figures/avg_med_rem_2015_2017.png')
 plt.clf()
 plt.close()
@@ -159,6 +161,7 @@ for i in range(0, len(plot_dfs)):
 	plot_dfs[i]['norm_'+plot_series[i]] = plot_dfs[i][plot_series[i]] / norm_factor
 	plt.plot(plot_dfs[i]['monthly_date'], plot_dfs[i]['norm_'+plot_series[i]], label = labels[i])
 plt.legend()
+plt.axvline(x = '2016-02', color = 'black', linestyle='-')
 plt.title("Normalized Monthly Remuneration 2015-2017")
 plt.savefig('/Users/Valerie/CBAs-Inflation-Git/drafts/tables-figures/normalized_avg_med_rem_2015_2017.png')
 plt.clf()
@@ -175,8 +178,9 @@ for i in range(0, len(plot_dfs)):
 	plot_dfs[i]['norm_'+plot_series[i]] = plot_dfs[i][plot_series[i]] / norm_factor
 	plt.plot(plot_dfs[i]['monthly_date'], plot_dfs[i]['norm_'+plot_series[i]], label = labels[i])
 plt.legend()
[…]

### 2021-04-26 Drafts/Draft_1.tex
Prospectus and Slides updates

@@
 Furthermore, we assess the degree of wage rigidity for firms that negotiate just before and just after the inflationary news shock.
-This research may have important implications for both traditional and expectations-based monetary policy: In particular, we can inferences about the optimal timing of monetary policy using our precise measure of the time-varying degree of nominal wage rigidity.
+This research may have important implications for both traditional and expectations-based monetary policy: In particular, we can make inferences about the optimal timing of monetary policy using our precise measure of the time-varying degree of nominal wage rigidity.
 Additionally, this research may highlight the importance of the collective bargaining channel of expectations-based monetary policy.
@@
 		\subsection{Main Research Questions}
-		In this project, we are mainly interested in estimating the effects of inflationary news shocks on collective bargaining outcomes in Brazil and understanding macroeconomic implications of these effects.
+		In this project, we are mainly interested in estimating the effect of the inflationary news shock on collective bargaining outcomes in Brazil and understanding macroeconomic implications].
 We summarize the main quesstions we will address below:
@@
 That is, nominal wages cannot be reduced unless the union approves the reduction first.
-Without an affirmation from the workers' union, if nominal wages fall, firms are legally required to offer workers additional hours.
 Conversely, a reduction in workers' hours must be compensated by a commensurate increase in the nominal wage rate.
 \cite{EnglishTranslationConsolidation2010}
-		Given the constraints facing firms, wages (measured in average monthly remuneration) may still be reduced through separations, or by reclassifying workers to occupations with lower nominal wage levels.
+
+		Given the constraints facing firms, wages (measured in average monthly remuneration) may still be reduced through three main channels: 
+		\begin{enumerate}
+			\item \textbf{Employment Protection Program} ``Programa de Proteção ao Emprego'' (PPE): This was implemented from July 7, 2015 -  December 31 to protect employment levels during the recession.
+During this period,  for a maximum of 12 months firms could opt into PPE to temporarily reduce hours by up to 30\% with a proportional reduction in monthly remuneration.
+The Brazilian government would compensate workers with 50\% of the wage reduction.
+In our dataset, CBAs involving firms joining PPE begin appearing in 2016, although uptake probably occurred as early as the program start date in July 2015.
+In future research, it may be interesting to compare wage trends for firms that opted into PPE with those that did not, as well as interactions with collective bargaining activity.
+			\item \textbf{separations}: Our measure of wages (average firm-level monthly remuneration) includes the monthly remunerations of workers who separate from the firm mid-month.
+In effect, mid-month separations would appear in the dataset as a reduction in monthly remuneration.
+In the future, we will analyze and compare results using a balanced panel of workers from the RAIS data to gauge the importance of the separations channel.
+			\item \textbf{reclassifying occupations}: Part of the reduction in average firm-level monthly remuneration may be due to workers being reclassified to occupations with lower implied nominal wages.
+Similar to the case of separations, reclassifications of this type would register as reductions in wages.
+		\end{enumerate}
 
@@
 We use the count of employed workers in a given month as the main employment variable of interest.
-		\begin{figure}
-			\centering
-			\caption{INSERT RAIS SUM STATS}
-			\label{tab:SumStatsRAIS}
-		\end{figure}
+		% \begin{figure}
+		% 	\centering
+		% 	\caption{INSERT RAIS SUM STATS}
+		% 	\label{tab:SumStatsRAIS}
+		% \end{figure}
 
-		\begin{figure}
-			\centering
-			\caption{INSERT RAIS RAW SAMPLE}
-			label{tab:RawDataRAIS}
-		\end{figure}
+		% \begin{figure}
+		% 	\centering
+		% 	\caption{INSERT RAIS RAW SAMPLE}
+		% 	label{tab:RawDataRAIS}
+		% \end{figure}
 
@@
 The key variables we use for this study the CBA start and end dates, and company names, which we use to link the CBA data with RAIS.
-		\begin{figure}
-			\centering
-			\caption{INSERT SM SUM STATS}
-			\label{tab:SumStatsRAIS}
-		\end{figure}
+		% \begin{figure}
+		% 	\centering
+		% 	\caption{INSERT SM SUM STATS}
+		% 	\label{tab:SumStatsRAIS}
+		% \end{figure}
 
-		\begin{figure}
-			\centering
-			\caption{INSERT SM RAW SAMPLE}
-			label{tab:RawDataRAIS}
-		\end{figure} 
-		\subsection{FGV IBGE}
-		We use aggregate consumer inflation expectations fron the Getuglio Vargas Foundation (FGV).
+		% \begin{figure}
+		% 	\centering
+		% 	\caption{INSERT SM RAW SAMPLE}
+		% 	label{tab:RawDataRAIS}
+		% \end{figure} 
+		\subsection{FGV IBRE}
+		We use aggregate consumer inflation expectations from the Getulio Vargas Foundation (FGV).
 In the future, we may gain access to firm-specific expectations of their own prices and costs.
 This information would alllow us to estimate the relationship between expectations and wage adjustments more precisely.
-		\begin{figure}
-			\centering
-			\caption{INSERT FGV SUM STATS}
-			\label{tab:SumStatsRAIS}
-		\end{figure}
+		% \begin{figure}
+		% 	\centering
+		% 	\caption{INSERT FGV SUM STATS}
+		% 	\label{tab:SumStatsRAIS}
+		% \end{figure}
 
-		\begin{figure}
-			\centering
-			\caption{INSERT FGV RAW SAMPLE}
-			label{tab:RawDataRAIS}
-		\end{figure}
+		% \begin{figure}
+		% 	\centering
+		% 	\caption{INSERT FGV RAW SAMPLE}
+		% 	label{tab:RawDataRAIS}
+		% \end{figure}
 
@@
 		We use the March 16 protests and anticipation of Rousseff's impeachment as the inflationary news shock.
-We estimate the effects on wages over time for firms in the pre- and post- shock groups.
-We restrict our sample to the subset of firms with contracts of at least 12 months durations beginning either in January 2016 or May 2016.
-We use January and May 2016 as our pre- and post-shock dates, respectively for two reasons: 1.) Average inflation expectations differ substantially across these periods\footnote{See Figure \ref{fig:Inflation}}, and 2.) most CBAs in a given year occur during these months.\footnote{See Figure \ref{fig:CBAProb}} This allows us to take advantage of a substantial part of the variation in the data.
+We estimate the effects on wages and number of workers over time for firms in the pre- and post- shock groups.
+We restrict our sample to the subset of firms with contracts beginning either in January 2016 or May 2016.
+We use January and May 2016 as our pre- and post-shock dates, respectively, for two reasons: 1.)the trajectoryies of average inflation expectations differ substantially across these periods, \footnote{See Figure \ref{fig:Inflation}} and 2.) most CBAs in a given year occur during these months.\footnote{See Figure \ref{fig:CBAProb}} This allows us to take advantage of a substantial part of the variation in the data.
 	\begin{align}
@@
 	In this specification, $D_{i,t}$ is a dummy variable that equals 1 if firm $i$ is in the post-shock group and if the time period of the observation is equal to $t$.
-A negative estimate for $\beta_t$ would suggest that at time period t, the log wages for firms in the post-shock group fell by more than the pre-shock group.
-$\alpha_i$ is a firm-level fixed effect, and $\delta_t$ is a time fixed effect.
+A negative estimate for $\beta_t$ would suggest that at time period t, the log wages for firms in the post-shock group fall by more than the pre-shock group.
+$\alpha_i$ is a firm-level fixed effect, and $\delta_t$ is a time fixed effect for each period in the sample.
 
-	If wages were perfectly flexible, our prediction, in line with Olivei and Tenreyro (2007), would be that the $\beta_t$s would be most negative just after the shock and attenuate to zero over time.
-However, given the presence of legally required downward nominal rigidity, firms are left with only indirect ways to reduce workers' monthly earnings.
-For example, monthly earnings would fall if firms separate from workers or reclassify them to occupations with lower implied wages.
-Note that simply reducing hours would not result in a reduction in monthly earnings, since the Brazilian Consolidation of Labor Laws (CLT) requires that hourly reductions be compensated with a commensurate increase in nominal wage rates.
-Moreover, because we restrict our analysis to a balanced panel, nominal wage reductions cannot be explained by firm exit/entry.
+	If wages were perfectly flexible, our prediction, in line with Olivei and Tenreyro (2007), would be that the $\beta_t$ estimates should be most negative just after the shock and attenuate to zero over time.
+However, for some firms, the presence of legally-required downward rigidity is mitigated by the provisional Employment Protection Program (PPE), which allowed participating firms to temporarily cut hours (and monthly remuneration) by up to 30\%.
+For firms that did not opt into PPE, a reduction in nominal monthly remuneration is feasible only through separations or by reclassifying workers.
+(Note that the sample is a balanced panel of firms, so nominal wage reductions cannot be explained by firm exit/entry.) In theory, firms that opted into PPE would use the additional wage flexibility temporarily afforded to behave in a manner more consistent with the predictions of Olivei and Tenyreyro, i.e., cut wages as inflation falls.
+Participation in PPE represents another form of treatment leading to differential wage flexibility across firms.
+This potential treatment effect merits study both independently and in connection to the timing of CBAs channel, which we focus on here.
 
@@
 			\label{fig:NormWages}
-			\includegraphics[scale = .65]{tables-figures/normalized_avg_med_rem_2015_2017.png}
+			\includegraphics[scale = .65]{tables-figures/normalized_avg_rem_2015_2017.png}
 			\floatfoot{Each series is normalized by the values by December 2015 values.
-The sample is restricted to a balanced panel of firms between 2015 - 2017 with contracts beginning either in January or May 2016 and lasting at least 12 months.}
+The sample is restricted to a balanced panel of firms between 2015 - 2017 with contracts beginning either in January or May 2016.
+See appendix for figure including median remuneration.}
 		\end{figure}
@@
 			\label{fig:NormEmp}
-			\includegraphics[scale = .65]{tables-figures/normalized_avg_med_emp_2015_2017.png}
+			\includegraphics[scale = .65]{tables-figures/normalized_avg_emp_2015_2017.png}
 			\floatfoot{Each series is normalized by the values by December 2015 values.
-The sample is restricted to a balanced panel of firms between 2015 - 2017 with contracts beginning either in January or May 2016 and lasting at least 12 months.}
+The sample is restricted to a balanced panel of firms between 2015 - 2017 with contracts beginning either in January or May 2016 See appendix for figure including median number of workers.}
 		\end{figure}
+		Next, we present results from the regressions the $\beta_t$ estimates from the difference in differences specifications in equations \ref{eqn:rem_DiD} and  \ref{eqn:emp_Did}.
+For this analysis, we restrict the sample to a balanced panel of firms between 2015 and 2016 whose contracts begin either in January 2016 or May 2016 and last at least 12 months.
+Below are the summary statistics for the dataset used in estimation.
 
+		\begin{table}[!ht]
+			\caption{Group Composition by Sector}
+			\label{tab:sectors}
+			\begin{tabular}{lrrr}
+				\input{tables-figures/sectors}
+			\end{tabular}
+		\end{table}
+
+	\input{tables-figures/DiD_summ_stats}[!ht]
 		In Figure \ref{fig:NormEmp}, we note that the average and median number of workers per firm is falling during the sample period.
@@
 
-		Figure \ref{fig:RegressionResults} plots the $\beta_t$ estimates from equations \ref{eqn:emp_Did} and \ref{eqn:rem_DiD} above.
 		 Each point represents the difference in nominal wage growth between firms that negotiated just before the March 2016 shock ($t=0$) and just after.
-We restrict the sample to a balanced panel of firms between 2015 and 2016 whose contracts begin either in January 2016 or May 2016 and last at least 12 months.
 		\begin{figure}[!ht]
 			\centering
-			\caption{Response of Wages and Employment Given CBA Timing}
-			\includegraphics[scale = .49]{tables-figures/DiD_Plots_Weighted_Employment_Rem.png}
+			\caption{Response of Nominal Wages and Employment Given CBA Timing}
+			\includegraphics[scale = .375]{tables-figures/DiD_Plots_Weighted_Employment_Rem.png}
 			\label{fig:RegressionResults}
@@
 The results for the wage regressions use analytic weights based on the number of workers at firm $i$ in period $t$.
-}
+Dashed lines represent when firms in the postshock group would have renegotiations if on a fixed 12-month schedule.}
 		\end{figure}
-		Contrary to our initial prediction, the DiD results suggest that firms that negotiated contracts just after disinflationary shock had higher nominal wage levels than their counterparts.
-Furthermore, this result is not accompanied by a relative reduction in the number of workers for the post-shock group.
+		Contrary to our initial prediction, the DiD results suggest that firms that negotiated contracts just after disinflationary shock had higher nominal wage levels than their counterparts, controlling for time- and firm-level fixed effects.
+Furthermore, this result is not accompanied by a relative reduction in the number of workers in the post-shock group.
 Instead, firms in the post-shock exhibit both relatively high nominal wages and employment levels, despite the fact that the real cost to group suggest that the difference in nominal wage growth is statistically indistinguishable from zero until January 2015 (t= -14), represented by the red dashed line on the graph.
@@
 	Tables \ref{tab:wage_reg} and \ref{tab:emp_reg} show several variants of the specifications displayed in Figure \ref{fig:RegressionResults}. 
-		\begin{figure}[!ht]
+		\begin{table}[!ht]
 				\caption{Log Wage Regressions}
@@
 				\floatfoot{B/ NB = Balanced/ Unbalanced, W/ NW = Weighted/ Unweighted. \\ Weights are based on number of workers at firm $i$ in period $t$.}
-		\end{figure}
+		\end{table}
 
-		\begin{figure}[!ht]
+		\begin{table}[!ht]
 			\caption{Log Employment Regressions}
@@
 			\floatfoot{B/ NB = Balanced/ Unbalanced}
-		\end{figure}
+		\end{table}
 
@@
 	\section{Future Research}
-	Further analysis is required to pin down what factors are driving the reduction in nominal wages across all firms during the 2015-2016 recession.\footnote{See Figure \ref{fig:NormWages}} One possible explanation is that firms separate from workers during the recession, curtailing the workers' monthly earnings.
+	Further analysis is required to pin down what factors are driving the reduction in nominal wages across all firms during the 2015-2016 recession.\footnote{See Figure \ref{fig:NormWages}} One possible explanation is that a large share of firms may be opting into PPE, as discussed above.
+Another possible explanation is that firms separate from workers at a higher rate during the recession, curtailing the workers' monthly earnings.
 There is some preliminary evidence of this in Figure \ref{fig:NormEmp}.
 Another potentially complementary explanation is that firms reclassify workers to occupations with lower implied nominal wages.
-While we have not yet carried out the analyses to corroborate whether the latter explanation is pertinent, we may have sufficient data to accomplish this in the near future.
+While we have not yet carried out the analyses to distinguish between these explanations, we may have sufficient data to accomplish this in the near future.
 
-	Another salient question is why firms that negotiate wages just after the shock set higher nominal wages than firms negotiating just before Figure \ref{fig:RegressionResults} One possible explanation is that firms in the post-shock group have higher levels of optimism about the future state of the economy and their own future profit margins, which may lead them to set higher wage levels.
+	Furthermore, the sheer size in the drop of nominal monthly remuneration during the recession across CBA groups suggests that it may be beneficial to broaden the scope our research beyond the inflationary news shock narrative to focus on the labor market effects of the recession as a whole.
+Relatedly, it may beneficial to reconsider whether the standard predictions of macro models (i.e., inflation and GDP are positively correlated) are applicable to the Brazilian context in 2015-2016.
+\cite{goncalvesAssessingBrazilianMacroeconomic2016a} find that during periods of high volatility in Brazil, inflation and output are negatively correlated.
+If this is true for the 2015-2016 recession, our predictions for the direction of wage and employment  variables in response to the disinflationary news may not necessarily align with the findings of Olivei and Tenreyro.
+An interesting question is what the model-based implications for the direction of wages and employment during a high volatility period would be, and whether these predictions hold empirically.
+
+	Another salient question is why firms that negotiate wages just after the shock set higher nominal wages than firms negotiating just before Figure \ref{fig:RegressionResults}.
+An important piece of this puzzle could be that the result is driven by particular sectors, or by potential cyclicality in the bargaining power of workers, particularly in the post-shock group.
+While we do not currently have evidence this is the case, further work is required to corroborate this empirically.
+Another possible explanation is that firms in the post-shock group have higher levels of optimism about the future state of the economy and their own future profit margins, which may lead them to set higher wage levels.
 We may gain access to microdata on firms' own-price and own-cost expectations that would allow us to study the link between individual firms' expectations and negotiated wages as well as hiring and firing decisions.
 
-	Following Olivei and Tenreyo (2007), we may adapt a dynamic stochastic general equilibrium model with uneven staggering of wage contracts and downward rigidity.
+	In a later stage of the paper, we plan to use a structural model to rationalize our empirical findings.
+For example, following Olivei and Tenreyo (2007), we may adapt a dynamic stochastic general equilibrium model with uneven staggering of wage contracts and downward rigidity.
+Another approach would be to adapt a DMP model to include inflation expectations.
+\cite{fukuiTheoryWageRigidity2020}
 
-	Another approach would be to adapt a DMP model to include inflation expectations. \cite{fukuiTheoryWageRigidity}
+\section{Appendix}
+	\begin{figure}[!ht]
+		\centering
+		\caption {Typical CBA in Sistema Mediador}
+		\label{typical_CBA}
+		\includegraphics[scale=.6]{tables-figures/Typical_CBA.png}
+		\floatfoot{Excerpt from \cite{lagosLaborMarketInstitutions2021}.}
+	\end{figure}
+	\begin{figure}[!ht]
+		\centering
+		\caption{Monthly Earnings per Worker}
+		\label{fig:NormWages_Med}
+		\includegraphics[scale = .65]{tables-figures/normalized_avg_med_rem_2015_2017.png}
+		\floatfoot{Each series is normalized by the values by December 2015 values.
+The sample is restricted to a balanced panel of firms between 2015 - 2017 with contracts beginning either in January or May 2016.}
+	\end{figure}
 
+	\begin{figure}[!ht]
+		\centering
+		\caption{Number of Workers}
+		\label{fig:NormEmp_Med}	
+		\includegraphics[scale = .65]{tables-figures/normalized_avg_med_emp_2015_2017.png}
+		\floatfoot{Each series is normalized by the values by December 2015 values.
+The sample is restricted to a balanced panel of firms between 2015 - 2017 with contracts beginning either in January or May 2016.}
+		\end{figure}
 % =================================================================================================

### 2021-04-26 Drafts/QE_Slides_April_2021.tex
Prospectus and Slides updates

\documentclass[aspectratio=169]{beamer}
\usetheme{Madrid}
\usepackage[citestyle = authoryear]{biblatex}
\bibliography{Bibliography-CBAs-Inflation}
\AtBeginBibliography{\scriptsize}
\usepackage{adjustbox}
\usepackage{color}
% \setbeamertemplate{footline}[frame number]
\title{Bringing Expectations to the Collective Bargaining Table}
\author{Valerie Boctor}
\subtitle{Evidence From Brazilian Firms}
\date{\today}
\usepackage{booktabs}
\begin{document}
	\begin{frame}
	\titlepage
	\end{frame}

	% \begin{frame}
	% \frametitle{Outline}
	% \tableofcontents
	% \end{frame}

	\section{Introduction}
	\begin{frame}
		\frametitle{Motivation}
		\begin{itemize}
			\item Understanding wages and employment during crisis in a high inflation environment
			\item Brazil as a paradigm of emerging economies
			\item Relevance to US: past and present
		\end{itemize}
	\end{frame}

	\begin{frame}
	\frametitle{Main Questions}
		\begin{itemize}
			\item How does the timing of CBAs affect the response of wages and employment to the inflationary news shock?
			\item To what extent do firm-level wages exhibit nominal rigidities and why?
			\item What are the implications of wage and employment behavior in response to the shock for expectations-based monetary policy?
		\end{itemize}
	\end{frame}

	\begin{frame}
	\frametitle{Economic Significance}
		\begin{itemize}
			\item Timing of monetary policy shocks (\cite{oliveiTimingMonetaryPolicy2007})
			\item Nominal rigidities \--- wage adjustment asymmetries (\cite{taylorStayingPowerStaggered2016})
			 \item Real wages \& response of employment (\cite{cardUnexpectedInflationReal1990})
			 \item Monetary policy application: inflation targeting
				\begin{itemize}
					\item Traditional channels are consumption and investment. Analyze potential CBA channel
				\end{itemize}
		\end{itemize}
	\end{frame}

	\section{Background}
		\subsection{The Inflationary News Shock}
			\begin{frame}
				\frametitle{Background I: The Inflationary News Shock}
				\centering
			 	\textbf{Key events: Dual economic and political crises in Brazil 2015 - 2016}
				\begin{enumerate}
					\item \underline{January 2015}: Rousseff begins 2nd term amid allegations of corruption, fiscal malfeasance. Economic recession begins.
					\item \underline{March 2016}: \textbf{Inflationary news shock} \---wave of protests solidifies calls for Rousseff's impeachment
					\item \underline{April 2016}: Rousseff is formally impeached
					\item \underline{May 2016}: Rousseff's powers suspended; V.P. Michel Temer takes over, appoints Ilan Goldfajn Brazilian Central Bank chief.
				\end{enumerate}
			\end{frame}

			\begin{frame}
				\frametitle{Background I: The Inflationary News Shock}
				\centering
				\includegraphics[scale=.3]{old_figures/goldfajn.png}
			\end{frame}

			\begin{frame}
				\frametitle{Background I: The Inflationary News Shock}
					\centering
					\includegraphics[scale=.35]{tables-figures/Realized_Inflation_and_Consumer_Expectations.png}
			\end{frame}

			\begin{frame}
				\frametitle{Background I: The Inflationary News Shock}
					\centering
					\includegraphics[scale=.55]{old_figures/IBRE_Consumer_Expectations}
			\end{frame}

			\begin{frame}
				\frametitle{Background I: The Inflationary News Shock}
					\centering
					\includegraphics[scale=.55]{old_figures/IBRE_Firm_Expectations}
			\end{frame}

		\subsection{CBAs in Brazil}		
			\begin{frame}
				\frametitle{Background II: CBAs in Brazil}
				\begin{itemize}
					\item ~65\% of formal sector workers are unionized in Brazil (compare with ~11\% in US).\cite{byjellevisserTrendsCollectiveBargaining2017}
					\item CBAs typically fix a \textbf{nominal wage level} for the duration of the contract. 
					\item \textbf{Brazilian Consolidation of Labor Laws}: Nominal wages cannot be reduced without workers' approval. 
					\item A reduction in hours must be paired with a commensurate increase in the hourly rate.
					\item Maximum duration of a CBA is 24 months (can be extended indefinitely).
					\item Agreements are either at the firm- or sector- level. We focus on firm-level agreements.
				\end{itemize}
			\end{frame}

			\begin{frame}
				\frametitle{Timing of CBAs in Brazil}
				\begin{center}
				\includegraphics[scale=.45]{tables-figures/cba_hazard.png}
				\end{center}
			\end{frame}

			\begin{frame}
				\frametitle{Timing of CBAs in Brazil}
				\centering
				\includegraphics[scale=.35]{tables-figures/cba_probability.png}
			\end{frame}

	\begin{frame}
		\frametitle{Key Takeaways}
		\begin{itemize}
			\item Large drop in inflation expectations coincides with Rousseff's impeachment and  replacement by Temer (and Goldfajn). This represents a plausible inflationary news shock
			\item Variation in the timing of CBAs is plausibly exogenous. 
		\end{itemize}
	\end{frame}

	\section{Data}
		\subsection{RAIS}
		\begin{frame}
			\frametitle{Data}
			\begin{itemize}
				\item \textbf{RAIS} \--- Matched employer-employee data 
				\begin{itemize}
					\item Key variables: Monthly remuneration, employment status, 6-digit occupation code, 2-digit sector code, firm ID, worker ID.
					\item Note: ``Monthly remuneration'' measured observed earnings per worker in a given month. We do not observe wage rates and hours separately.
				\end{itemize}
			\end{itemize}
		\end{frame}

		\subsection{Sistema Mediador}
			\begin{frame}
				\frametitle{Data}
				\label{sistema_mediador}
				\begin{itemize}
					\item \textbf{RAIS} \--- Matched employer-employee data 
					\begin{itemize}
						\item Key variables: Monthly remuneration, employment status, 6-digit occupation code, 2-digit sector code, firm ID, worker ID.
						\item Note: ``Monthly remuneration'' measured observed earnings per worker in a given month. We do not observe wage rates and hours separately.
					\end{itemize}
					\item \textbf{Sistema Mediador} \--- collective bargaining agreements (single firm - employee contracts)
			 		\begin{itemize}
				 		\item Key variables: company names, start/end dates of contracts
					 \end{itemize}
				 \end{itemize}
				 \hyperlink{typical_cba}{\beamerbutton{Example of CBA Contract}}
			 \end{frame}

		\subsection{FGV IBRE}	
			\begin{frame}
				\frametitle{Data}
				\begin{itemize}
					\item \textbf{RAIS} \--- Matched employer-employee data 
					\begin{itemize}
						\item Key variables: Monthly remuneration, employment status, 6-digit occupation code, 2-digit sector code, firm ID, worker ID.
						\item Note: ``Monthly remuneration'' measured observed earnings per worker in a given month. We do not observe wage rates and hours separately.
					\end{itemize}
					\item \textbf{Sistema Mediador} \--- collective bargaining agreements (single firm - employee contracts)
			 		\begin{itemize}
				 		\item Key variables: company names, start/end dates of contracts
					 \end{itemize}

				 \hyperlink{typical_CBA}{\beamerbutton{Example of CBA Contract}}
					\item \textbf{FGV IBRE}\--- aggregate inflation expectations	(May gain access to firm-level price and cost expectations)
					\begin{itemize}
						\item Key variables: Aggregate inflation expectations, firm and consumer confidence indices. 
						\item Inflation expectations capture households' projections for inflation over the next 12 months.
					\end{itemize}
				\end{itemize}
			\end{frame}

	\section{Preliminary Findings}

		% \begin{frame}
		% 	\frametitle{Preliminary Results: Real Monthly Remuneration}
		% 	\label{normalized_real_avg_rem}
		% 	% \hyperlink{normalized_a vg_med_rem}{\beamerbutton{Median Plots}}
		% 	\centering
		% 	\includegraphics[scale = .4]{tables-figures/normalized_real_avg_rem_2015_2017.png}
		% \end{frame}

		\begin{frame}
			\frametitle{Preliminary Results: Nominal Monthly Remuneration}
			\label{normalized_avg_rem}
			\hyperlink{normalized_avg_med_rem}{\beamerbutton{Median Plots}}
			\centering
			\includegraphics[scale = .4]{tables-figures/normalized_avg_rem_2015_2017.png}
		\end{frame}

		\begin{frame}
			\frametitle{Preliminary Results: Number of Workers}
			\label{normalized_avg_emp}
			\hyperlink{normalized_avg_med_emp}{\beamerbutton{Median Plots}}
			\centering
			\includegraphics[scale = .4]{tables-figures/normalized_avg_emp_2015_2017.png}
		\end{frame}

		\begin{frame}
			\frametitle{Reduced-Form Model}
		 	Using a DiD approach, we measure effect on CBA timing relative to the shock on nominal wage and employment levels.
			\begin{align}
				\intertext{We write down the reduced form model for wages as:} \label{eqn:rem_DiD}
				\log{w_{i,t}} & = \alpha_i + \delta_t + \sum_{t=\text{Jan 2015}, t\neq t^{\text{shock}}}^{\text{Dec 2017}} \beta_{t}D_{i,t} + \varepsilon_{i,t} 
				\intertext{Similarly, the model for number of workers is:}  \label{eqn:emp_Did}
				\log{n_{i,t}} & = \alpha_i + \delta_t + \sum_{t=\text{Jan 2015}, t\neq t^{\text{shock}}}^{\text{Dec 2017}} \beta_{t}D_{i,t} + \varepsilon_{i,t} 
			\end{align}

			\begin{itemize}
				\item balanced panel of firms from 2015 - 2017 
				\item contracts set either in Jan 2016 (pre-shock) or May 2016 (post-shock)
			\end{itemize}
		\end{frame}

		\begin{frame}
			\scriptsize{\input{tables-figures/DiD_summ_stats}}
		\end{frame}

		\begin{frame}
			\centering
			\begin{table}
				\scriptsize{\begin{tabular}{lrrr}
				\input{tables-figures/sectors}
				\end{tabular}
			}
			\end{table}
		\end{frame}

		\begin{frame}
			\frametitle{DiD Results}
				\centering
				\includegraphics[scale = .225]{tables-figures/DiD_Plots_Weighted_Employment_Rem.png}
		\end{frame}

		\begin{frame}
			\frametitle{Preliminary Findings}
			\begin{itemize}
				\item For both groups of firms, large, persistent drop in nominal wages between Dec. 2015 and Jan. 2016, despite legally-required downward wage rigidity.
				\item This may be driven by mid-month separations, and/or reclassification of workers.
				\item Contrary to prediction, firms with CBAs post-shock actually had higher nominal wages than firms with pre-shock CBAs. 
				\item Seasonality in coefficient estimates
			\end{itemize}
		\end{frame}

			\begin{frame}
			\frametitle{Preliminary Findings}
			\begin{itemize}
				\item For both groups of firms, large, persistent drop in nominal wages between Dec. 2015 and Jan. 2016, despite legally-required downward wage rigidity.
				\item This may be driven by mid-month separations, and/or reclassification of workers.
				\item Contrary to prediction, firms with CBAs post-shock actually had higher nominal wages than firms with pre-shock CBAs. 
				\item Seasonality in coefficient estimates
			\end{itemize}
			{\color{red}{Possible explanations: optimism, sectoral heterogeneity, cyclicality in workers' bargaining power, regime switching. }}
		\end{frame}

		\section{Next Steps \& Future Research}
		\begin{frame}
			\frametitle{Next Steps \& Future Research}
			\begin{itemize}
				\item Broaden scope of research question; move beyond inflation shock narrative
				\item Explore heterogeneity across sector
				\item Unpack key drivers in observed nominal wage reduction across groups: separations, reclassifications, or both? 
				\item Estimate effects for real wages and compare with nominal wage and number of workers results.
				\item Pending data access, explore ``optimism'' narrative using own-price/cost expectations microdata. 
				\item Analyze firms' other margins in response to the shock. Data sources?
			\end{itemize}
		\end{frame}
		\begin{frame}
		\frametitle{Structural Identification Strategy}
			\begin{itemize}
				\item Adapt a DSGE model with uneven wage staggering wage model with collective bargaining to assess impacts of monetary policy (\cite{oliveiTimingMonetaryPolicy2007}) or;
				\item Adapt a DMP model with inflation expectations to understand implications for aggregate wages and employment.
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{References}
			\printbibliography
		\end{frame}

	\section{Additional Projects}
		\begin{frame}
			\centering
			\Huge{Additional Projects}
		\end{frame}

		\begin{frame}
			\frametitle{Regional Inflation Expectations During Covid \& The Phillips Curve}
			Main questions: 
			\begin{itemize}
				\item How do regional economic conditions \& household characteristics (e.g. employment status) during affect household inflation expectations?
				\item What is the slope of the Phillips curve, measured using the regional inflation expectations data?
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{Regional Inflation Expectations During Covid \& The Phillips Curve}
			\centering
			\includegraphics[scale=.35]{tables-figures/Pre-Covid.png}
		\end{frame}

		\begin{frame}
			\frametitle{Regional Inflation Expectations During Covid \& The Phillips Curve}
			\centering
			\includegraphics[scale=.35]{tables-figures/Post-Covid.png}
		\end{frame}

		\begin{frame}
			\frametitle{Household Wealth Perceptions and Labor Supply Decisions}
			\begin{itemize}
				\item \textbf{Main question:} How do households' wealth perceptions affect 
				\item i.e., is there a ``Behavioral'' Laffer Curve?
				\item Potential data source: DNB Household Survey
				\item Ideal setting: Look for tax reform as source of exogenous variation in wealth perceptions.
			\end{itemize}
		\end{frame}

		\begin{frame}
			\centering
			\Huge{Appendix}
		\end{frame}

		\begin{frame}
			 	\frametitle{Sistema Mediador: Typical CBA (excerpt from \cite{lagosLaborMarketInstitutions2021})}
			 	\label{typical_CBA}
			 	\tiny{Excerpt from (\cite{lagosLaborMarketInstitutions2021})}
				\centering
				\includegraphics[scale= .3]{tables-figures/Typical_CBA.png}
				\hyperlink{sistema_mediador}{\beamerbutton{Go Back}}
			 \end{frame}

		\begin{frame}
			\frametitle{Preliminary Results: Nominal Monthly Remuneration}
				\label{normalized_avg_med_rem}
				\hyperlink{normalized_avg_rem}{\beamerbutton{Go Back}}
				\centering
				\includegraphics[scale = .4]{tables-figures/normalized_avg_med_rem_2015_2017.png}
		\end{frame}

		\begin{frame}
			\frametitle{Preliminary Results: Number of Workers}
			\label{normalized_avg_med_emp}
			\hyperlink{normalized_avg_emp}{\beamerbutton{Go Back}}
			\centering
			\includegraphics[scale = .4]{tables-figures/normalized_avg_med_emp_2015_2017.png}
		\end{frame}

	\end{document}

### 2021-04-26 python/python_rev_2021/01_summary_stats.py

@@ -23,7 +23,20 @@ for i in date_vars:
 # np.sum(cba_df.COLUMN_NAME.count())
 
 # make a dummy variable that equals the sum of the month of contract and the contract duration is >= 12
-graph_df = graph_df[['identificad', 'monthly_date', 'rem','duration', 'cba_month', 'num_workers']]
+graph_df = graph_df[['identificad', 'monthly_date', 'rem', 'real_earnings','duration', 'cba_month', 'num_workers', 'clascnae20']]
+graph_df = graph_df.set_index('monthly_date')
+graph_df = graph_df.loc['2015-01':'2017-12']
+graph_df = graph_df.reset_index()
+balanced_ids = graph_df.groupby('identificad')['monthly_date'].nunique()
+balanced_ids = balanced_ids.reset_index()
+balanced_ids = balanced_ids.rename(columns = {'monthly_date':'num_months'})
+graph_df = balanced_ids.merge(graph_df, on = 'identificad', how = 'inner')
+graph_df = graph_df.set_index('monthly_date')
+
+graph_df = graph_df[graph_df.num_months == 36] # balanced panel!
+# graph_df = graph_df[graph_df.duration >= 12]
+
+graph_df = graph_df.reset_index()
 graph_df['cba_time'] = graph_df.monthly_date[graph_df.cba_month ==1]
 cba_jan16 = (graph_df.cba_time == '2016-01').astype('int')
 cba_may16 = (graph_df.cba_time== '2016-05').astype('int')
@@ -48,26 +61,23 @@ graph_df = cba_jan16_y_n.merge(graph_df, on = 'identificad',how = 'inner')
 graph_df= graph_df.rename(columns={"cba_may16_x": "cba_may16_y_n", "cba_may16_y": "cba_may16","cba_jan16_x":"cba_jan16_y_n", "cba_jan16_y": "cba_jan16"})
 
 graph_df = graph_df[graph_df.cba_may16_y_n != graph_df.cba_jan16_y_n]
-graph_df = graph_df.set_index('monthly_date')
-graph_df = graph_df.loc['2015-01':'2017-12']
-graph_df = graph_df.reset_index()
-balanced_ids = graph_df.groupby('identificad')['monthly_date'].nunique()
-balanced_ids = balanced_ids.reset_index()
-balanced_ids = balanced_ids.rename(columns = {'monthly_date':'num_months'})
[…]
